---
title: "Case Study 1: EDA to Create a new Beer for U.S. Craft Enthusiasts"
author: "Domicia Herring & Taylor Bonar"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
### a.	The introduction needs to be written as if you are presenting the work to the CEO and CFO of Budweiser (your client) and that they only have had one class in statistics.  If it sounds like a student presentation, that is not acceptable.  You may assume that the CEO and CFO gave you the data and gave you the directive to report any interesting finding that you may uncover through your analysis.

### b.	Briefly explain the purpose of the code. The explanations should appear as a sentence or two before or after the code chunk. Even though you will not be hiding the code chunks (so that I can see the code), you need to assume that the client can’t see them.
## Libraries & Package Setup
```{r}
# Load APIs
library(ggplot2)
library(dplyr)
library(maps)
library(mapproj)
library(stringr)
library(tidyr)
```

## Global Variable Setup & Cleanup
```{r}
# Importing input data
breweries = read.csv("Breweries.csv")
beers = read.csv("Beers.csv")

# Cleaning Data of invalid beginning space
breweries$State <- str_replace_all(breweries$State, "(\\s+)", "")
```

## Analysis

### How many breweries are present in each state?
```{r}
# Data tidying
## Remove DC as its not a state but capitol
breweriesStates = filter(breweries,!grepl('DC', State))

# Data Transformation
## Create data frame with State name and abbreviation
stateLookup = data.frame(abb = state.abb, stateName = state.name)

## Merge stateLookup with breweriesMapData
breweriesStates2 = merge(breweriesStates,stateLookup,by.x="State",by.y="abb")

## count each occurrence of breweries in each state
breweriesMapData = count(breweriesStates2,stateName)
## remap columns for future merge of states and map
colnames(breweriesMapData) = c("region","Breweries")
breweriesMapData$region <- tolower(breweriesMapData$region)

## Create merge between states lat/long and breweries map data
states <- map_data("state")
map.df <- merge(states, breweriesMapData, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]

## Create U.S. heat map to visualize states breweries count
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=Breweries))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Breweries by State")+coord_map()
```

### Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r}
# Join beer and breweries by common brewery identifier
beerAndBreweries = merge(beers, breweries, by.x="Brewery_id", by.y="Brew_ID")

# Remap columns named "Name" to be more descriptive/ubiquitous 
colnames(beerAndBreweries)[2] = "Beer_Name"
colnames(beerAndBreweries)[8] = "Brewery_Name"

# Show first six rows
head(beerAndBreweries)
# Show last six rows
tail(beerAndBreweries)
```

### For metrics and later data analysis, beers with missing IBUs and ABV were removed for the below reasons.
#### International bitterness units (IBUs) may not be included for all beers as some are not geared for their bitter flavors (e.g. lagers), but rather, IBU tends to be used as a quality control metric. IBUs tend not to be centered on marketing to impact a consumer's experience, but are simply there to satisfy some curiosities.
#### Alcohol by Volume (ABV) are not always printed, or rather required, to be displayed on beers by federal regulations. Instead, states regulate the advertisment of alcohol contents after the a unanimous vote in the Rubin vs Coors Brewing Co. 154 U.S.476 (1995). This court case argued against a 1935 decision ban on listing alcohol content to prevent "brewing wars".
```{r}
# Observe summary of data, only IBU and ABV contains NAs
summary(beerAndBreweries)

# We'll drop beers from the merged data that do not contain a value (i.e. NA)
beerAndBreweries2 = beerAndBreweries %>% drop_na(IBU, ABV)

# Observe data contains no NA values
summary(beerAndBreweries2)
```

### Exploring the median alcohol content and international bitterness unit for each state.
```{r}
# Calculate median of ABV or IBU by state grouping
medianABVbyState = aggregate(beerAndBreweries2$ABV,
          list(beerAndBreweries2$State),
          median)
medianIBUbyState = aggregate(beerAndBreweries2$IBU,
          list(beerAndBreweries2$State),
          median)

# Remap column names to be more descriptive
colnames(medianABVbyState) = c("State", "ABV")
colnames(medianIBUbyState) = c("State", "IBU")

# TODO: Add better visualization, too many states creates clutter and can overwhelm data visualization
ggplot(medianABVbyState, aes(x=State,y=ABV)) + 
  geom_col(fill = "#f28e1c") +
  labs(title = "Median Alcohol by Volume (ABV) for Each State") +
  theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1))

ggplot(medianIBUbyState, aes(x=State,y=IBU)) + 
  geom_col(fill = "#f28e1c") + 
  labs(title = "Median International Bitterness Units (IBU) for Each State") +
  theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1))
```

### Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
#Arrange to see in desc order by ABV
beerAndBreweries2 %>% group_by(State) %>% summarize(maxABV = max(ABV), count = n()) %>% arrange(desc(maxABV)) %>% print(n = 1)
#Arrange to see in order by IBU
beerAndBreweries2 %>% group_by(State) %>% summarize(maxIBU = max(IBU), count = n()) %>% arrange(desc(maxIBU)) %>% print(n = 1)

# find max of ABV, then show ABV level and State from the row
beerAndBreweries2[which.max(beerAndBreweries2$ABV),c(4,10)]
# find max of IBU, then show IBU and State from the row
beerAndBreweries2[which.max(beerAndBreweries2$IBU),c(5,10)]
```

### Summary statistics and distribution of the ABV variable.
```{r}
# Create initial summary of ABV content to gauge initial data metrics
summary(beerAndBreweries2$ABV)

# Compute the number of breaks by classes for histogram
## range = min is 0.027 and max is 0.125
## nclass.FD computes the number of classes for a histogram
breaks <- pretty(range(beerAndBreweries2$ABV), n = nclass.FD(beerAndBreweries2$ABV), min.n = 1)

# Compute binwidth for histogram based on suggested number of classes within the range of ABV
## (e.g. 0.03-0.025 = 0.005)
bwidth <- breaks[2]-breaks[1]

# Create histogram with spaced bins of ABV vs. number of beers
ggplot(beerAndBreweries2, aes(ABV)) +
  geom_histogram(fill = "#f28e1c", color="brown", binwidth = bwidth) +
  geom_density(alpha=.2, fill="blue") +
  labs(title="Distribution of Alcohol by Volume") +
  xlab("Alcohol by Volume (ABV)") +
  ylab("Number of Beers")

## TODO: Comment on statistics/distribution of ABV variable

```

### Is there an apparent relationship between the bitterness of the beer and its alcoholic content?
```{r}
# Create overall scatterplot to observe any clustering effect
beerAndBreweries2 %>% ggplot(aes(x=ABV, y=IBU)) + geom_point(colour = "red") + ggtitle("ABV vs. IBU") +
  xlab("ABV") + ylab("IBU")

# Create second overall scatterplot with trend line to observe relationship
ggplot(beerAndBreweries2, aes(x=ABV, y=IBU)) +
  geom_point(color = "#f28e1c", shape = "diamond", size = 1.5) +
  geom_smooth(method = 'gam', formula = y ~ s(x, bs = "cs")) +
  labs(title = "Relationship Between ABV and IBU", 
       subtitle = "An Upward Trend")

## TODO: Comment on relationship and elaborate
```

### Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.
```{r}
#Filter for all styles that contain 'Ale' and create a new data frame.
Ales <- beerAndBreweries2 %>% select(Style,ABV,IBU) %>%filter(stringr::str_detect(Style, 'Ale')) 

#Filter for all styles that contain 'India Pale Ale' and create a new data frame.
IndiaAles <- Ales %>%filter(stringr::str_detect(Style, 'India Pale Ale')) 
```

#### In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

### Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence. 
